---
title: "R_Recap_02"
author: "Felix Mueller"
date: "2/27/2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Prepare your R session
<div class="blue">
```{r Preparations, echo = T, warning =FALSE,message = F}
setwd("~/Google_Drive/github/R/R_Recap/session_02")
library(dplyr)
library(ggplot2)
```
</div>
# Iris dataset {.smaller}
<div class="blue">
1. sepal length in cm 
2. sepal width in cm 
3. petal length in cm 
4. petal width in cm 
5. class: 
-- Iris Setosa 
-- Iris Versicolour 
-- Iris Virginica
</div>
Source: https://archive.ics.uci.edu/ml/datasets/Iris

# Loading the Iris dataset {.smaller}
<div class="blue">
```{r iris, echo = T, warning =FALSE,message = F}
glimpse(iris)
```
</div>
# Boston Housing data {.smaller}
<div class="blue">
1. CRIM      per capita crime rate by town
2. ZN        proportion of residential land zoned for lots over 25,000 sq.ft.
3. INDUS     proportion of non-retail business acres per town
4. CHAS      Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)
5. NOX       nitric oxides concentration (parts per 10 million)
6. RM        average number of rooms per dwelling
7. AGE       proportion of owner-occupied units built prior to 1940
8. DIS       weighted distances to five Boston employment centres
9. RAD       index of accessibility to radial highways
10. TAX      full-value property-tax rate per $10,000
11. PTRATIO  pupil-teacher ratio by town
12. B        1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
13. LSTAT    % lower status of the population
14. MEDV     Median value of owner-occupied homes in $1000's
</div>
Source: http://archive.ics.uci.edu/ml/datasets/Housing

# Loading the Boston dataset {.smaller}
<div class="blue">
```{r housing data, echo = T, warning =FALSE,message = F}
library(MASS)
str(Boston)
```
</div>
