---
title: "R_Recap_02"
author: "Felix Mueller"
date: "2/27/2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Slide with Bullets

1. CRIM      per capita crime rate by town
2. ZN        proportion of residential land zoned for lots over 25,000 sq.ft.
3. INDUS     proportion of non-retail business acres per town
4. CHAS      Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)
5. NOX       nitric oxides concentration (parts per 10 million)
6. RM        average number of rooms per dwelling
7. AGE       proportion of owner-occupied units built prior to 1940
8. DIS       weighted distances to five Boston employment centres
9. RAD       index of accessibility to radial highways
10. TAX      full-value property-tax rate per $10,000
11. PTRATIO  pupil-teacher ratio by town
12. B        1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
13. LSTAT    % lower status of the population
14. MEDV     Median value of owner-occupied homes in $1000's

## Slide with R Output

```{r housing data}
setwd("~/Google Drive/github/R/R_Recap/session_02")
require(ggplot2)
require(dplyr)
```

```{r housing data}
data = read.table("housing.data.txt")
colnames(data) = c("CRIM","ZN","INDUS","CHAS","NOX","RM","AGE","DIS","RAD","TAX","PTRATIO","B","LSTAT","MEDV")
```

```{r iris}
glimpse(iris)
```
